---
title: "Homework 6"
author: "Xiaoni Xu"
date: "2024-11-25"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(ggplot2)

set.seed(1)
```


## Problem 1

Loading the weather data.
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities, \( \hat{R}^2 \) and \( \log(\hat{\beta}_0 \hat{\beta}_1) \)

```{r}
# Bootstrap function using broom:: functions
bootstrap_regression <- function(data, n_bootstrap = 5000) {
  n <- nrow(data)
  
  # Initialize vectors to store results
  r_squared <- numeric(n_bootstrap)
  log_beta_product <- numeric(n_bootstrap)
  
  for (i in 1:n_bootstrap) {
    bootstrap_sample <- data[sample(1:n, n, replace = TRUE), ]
    
    lm_fit <- lm(tmax ~ tmin, data = bootstrap_sample)
    
    r_squared[i] <- broom::glance(lm_fit)$r.squared
    
    # Extract coefficients using broom::tidy() and compute log(beta_0 * beta_1)
    coef_values <- broom::tidy(lm_fit) %>%
      filter(term %in% c("(Intercept)", "tmin")) %>%
      pull(estimate)
    
    beta_0 <- coef_values[1]
    beta_1 <- coef_values[2]
    log_beta_product[i] <- log(abs(beta_0 * beta_1)) # Use abs() to avoid log of negatives
  }
  
  # Return a data frame with the results
  tibble(
    r_squared = r_squared,
    log_beta_product = log_beta_product
  )
}

bootstrap_results <- bootstrap_regression(weather_df)

summary(bootstrap_results)
```


Plot the distribution of your estimates, and describe these in words.
 
```{r}
# Visualize the distribution of R-hat squared
ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "#007bbb", alpha = 0.7) +
  labs(
    title = "Distribution of R-hat Squared (\u005cR^2)",
    x = expression(hat(R)^2),
    y = "Frequency"
  ) +
  theme_minimal()

# Visualize the distribution of log(beta_0 * beta_1)
ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram(bins = 30, fill = "#79ba00", alpha = 0.7) +
  labs(
    title = "Distribution of log(beta_0 hat * beta_1 hat)",
    x = expression(log(hat(beta)[0] * hat(beta)[1])),
    y = "Frequency"
  ) +
  theme_minimal()
```
 
The distribution of \( \hat{R}^2 \) is unimodal and follows an approximate normal distribution, with most values clustering around the average \( R^2 \) = 0.9114 of the linear regression model.

The distribution of \( \log(\hat{\beta}_0 \hat{\beta}_1) \) is unimodal and follows an approximate normal distribution, with most values clustering around the average \( \log(\hat{\beta}_0 \hat{\beta}_1) \) = 2.013 of the linear regression model.



Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for both quantities.

```{r}
ci_r_squared <- quantile(bootstrap_results$r_squared, c(0.025, 0.975))
ci_log_beta_product <- quantile(bootstrap_results$log_beta_product, c(0.025, 0.975))

list(
  r_squared_ci = ci_r_squared,
  log_beta_product_ci = ci_log_beta_product
)
```
 

## Problem 2

Loading the homicide data
```{r}
hmc <- read_csv("data/homicide-data.csv")
```

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. 



For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}

```


For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}

```


Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.
```{r}

```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}

```


## Problem 3

Loading birthweight data; clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}

```


Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and 

```{r}

```


show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}

```


Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}

```



